---
output:
  rmarkdown::github_document:
    number_sections: false
    toc: false
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(knitr)
library(bscui)
library(base64enc)
library(htmltools)
library(dplyr)
library(xml2)
cranRef <- function(x){
  sprintf(
    "[%s](https://CRAN.R-project.org/package=%s): %s",
    x, x, packageDescription(x)$Title
  )
}
logo64 <- dataURI(file=here("supp/logo/bscui-medium.png"))
```

# bscui `r img(src=logo64, alt="logo", align="right", width="120px")`

<!--
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/bscui)](https://cran.r-project.org/package=bsui)
[![](http://cranlogs.r-pkg.org/badges/bsui)](https://cran.r-project.org/package=bsui)
-->

Produce interactive SVG figures with selectable and clickable user
interface elements. These figures can be integrated in markdown documents or
in shiny applications reacting to events triggered on them.
Features also include moving, zoom in/out and export in SVG or PNG formats.

## Installation

<!--

## From CRAN

```{r, eval=FALSE}
install.packages("bscui")
```

-->

### Dependencies

The following R packages available on CRAN are required:

```{r, echo=FALSE, results='asis'}
deps <- desc::desc_get_deps()
sdeps <- filter(deps, type %in% c("Depends", "Imports") & package!="R")
for(p in sdeps$package){
   cat(paste("-", cranRef(p)), sep="\n")
}
```

And those are suggested:

```{r, echo=FALSE, results='asis'}
wdeps <- filter(deps, type=="Suggests" & package!="R")
for(p in wdeps$package){
   cat(paste("-", cranRef(p)), sep="\n")
}
```

### From github

```{r, eval=FALSE}
devtools::install_github("patzaw/bscui")
```

## Usage

- Interactive figures: link to vignette
- Shiny applications: links (webR?)
