---
title: "Preparing SVG: examples, tips and tricks"
package: "bscui (version `r packageVersion('bscui')`)"
output:
   rmarkdown::html_document:
      number_sections: yes
      self_contained: true
      theme: cerulean
      toc: yes
      toc_float: yes
      fig_width: 7
      fig_height: 5
editor_options:
  chunk_output_type: console
---


```{r setup, message=FALSE, echo=FALSE, include=FALSE, cache=FALSE}
library(knitr)
opts_chunk$set(
   include=TRUE,
   echo=TRUE,
   message=TRUE,
   warning=TRUE,
   cache=FALSE,
   cache.lazy=FALSE
)
library(here)
library(bscui)
library(xml2)
library(dplyr)
library(readr)
##
data_dir <- here("inst/examples")
```

```{js, echo=FALSE}
function toggleCollapse(id) {
   var content = document.querySelector('#'+id);
   if(content.style.display === 'none' || content.style.display === ''){
      content.style.display = 'block';
   }else{
      content.style.display = 'none';
   }
}
```

```{css, echo=FALSE}
/* Hide the content of the collapsible div by default */
.collapsible {
   display: none;
   border: solid 1px black;
   background-color: #FAFA000F;
   margin: 5px 0px 5px 0px;
   padding: 5px;
}

.collapse-button {
   border: solid 1px black;
   border-radius: 3px;
   background-color: #CFCFCF;
}
```

This document describes how the SVG and related information used in the examples
of the 'bscui' package were prepared.

# Requirements

The following packages were used:

```{r}
library(bscui)
library(xml2)
library(dplyr)
library(readr)
library(stringr)
```

<!-- Collapsible sessionInfo -->
<button class="collapse-button" onclick="toggleCollapse('sessionInfo')">
Display session info
</button>

::: {#sessionInfo class="collapsible" style="margin-top:10px; margin-bottom:10px;"}

Built on `r Sys.Date()` 

```{r, class.source='fold-hide'}
sessionInfo()
```
:::

# Reading EBI anatomograms


## Original SVG

Anatomical diagrams are taken from
the [EBI gene expression
group](https://github.com/ebi-gene-expression-group/anatomogram).

```{r}
file_name <- "homo_sapiens.female.svg"
svg <- read_xml(file.path(
   "https://raw.githubusercontent.com",
   "ebi-gene-expression-group/anatomogram",
   "master/src/svg",
   file_name
))
```

## Extract element titles from the SVG

Organ identifiers and names were extracted directly from the SVG.

```{r}
get_element_titles <- function(x){
   id <- xml_attr(x, "id")
   children <- xml_children(x)
   children_titles <- tibble(id=character(), label=character())
   label <- character()
   if(length(children) > 0) for(i in 1:length(children)){
      child <- children[[i]]
      if(xml_name(child) == "title"){
         label <- xml_attr(child, "id")
      }else{
         children_titles <- children_titles |>
            bind_rows(get_element_titles(child))
      }
   }
   toRet <- tibble(id = id, label = label)  |>
      bind_rows(children_titles) |>
      filter(!is.na(id))
   return(toRet)
}
element_titles <- get_element_titles(svg)
```

```{r}
head(element_titles)
```

## Remove title elements from SVG

In order to avoid double title display, '<title>' elements were removed from the
original SVG.

```{r}
namespace_element_nodes <- xml_find_all(
   svg, "//namespace::*[name()='']/parent::*"
)
xml_attr(namespace_element_nodes, "xmlns") <- NULL
titles <- xml_find_all(svg, "//title")
for(to_remove in titles){
   xml_remove(to_remove)
}
xml_attr(namespace_element_nodes[1], "xmlns") <- "http://www.w3.org/2000/svg"
```

## Move organs in front

Finally, by choice, the organs were put in front of the body outline.

```{r}
children <- xml_children(svg)
xml_attr(namespace_element_nodes, "xmlns") <- NULL
to_move <- children[[which(xml_attr(children, "id") == "LAYER_EFO")]]
xml_remove(to_move)
xml_add_child(svg, to_move)
```

Here is the pre-processed SVG in a bscui figure:

```{r}
bscui(svg)
```

```{r, echo=FALSE}
write_xml(svg, file.path(data_dir, str_replace(file_name, ".svg$", ".svg.gz")))
write_tsv(
   element_titles,
   file.path(data_dir, str_replace(file_name, ".svg$", ".txt.gz"))
)
```

# Reading swissbiopics

## Read original SVG

The figure of animal cells was taken
from [SwissBioPics](https://www.swissbiopics.org/name/Animal_cell).

```{r}
file_name <- "Animal_cells.svg"
svg <- read_xml(file.path(
   "https://www.swissbiopics.org/",
   "api/image",
   file_name
))
```

## Remove text elements from SVG

Hidden text elements in the original SVG have consequence on the size and on
the scalling of the bscui figure. Therefore, they were removed from the SVG.

```{r}
namespace_element_nodes <- xml_find_all(
   svg, "//namespace::*[name()='']/parent::*"
)
xml_attr(namespace_element_nodes, "xmlns") <- NULL
titles <- xml_find_all(svg, "//text")
for(to_remove in titles){
   xml_remove(to_remove)
}
xml_attr(namespace_element_nodes[1], "xmlns") <- "http://www.w3.org/2000/svg"
```

Here is the pre-processed SVG in a bscui figure:

```{r}
bscui(svg)
```

```{r, echo=FALSE}
write_xml(svg, file.path(data_dir, str_replace(file_name, ".svg$", ".svg.gz")))
```

## Get cellular location information

Information about the different part of a cell were taken
from [UniProt](https://www.uniprot.org/locations).

```{r}
tmp_file <- tempfile(fileext=".txt.gz")
info <- download.file(
   url = "https://rest.uniprot.org/locations/stream?compressed=true&fields=id%2Cname%2Ccategory%2Cdefinition%2Ccontent%2Cgene_ontologies&format=tsv&query=%28*%29",
   destfile = tmp_file
)
info <- readr::read_tsv(
   tmp_file,
   col_types = strrep("c", 6)
)
```

```{r, echo=FALSE}
file_name <- "uniprot_cellular_locations.txt.gz"
readr::write_tsv(info, file=file.path(data_dir, file_name))
```

# Read wikipathways

## Read original SVG

Biological pathway SVGs were taken
from [WikiPathways](https://www.wikipathways.org/).

```{r}
pid <- "WP112"
file_name <- paste0(pid, ".svg")
svg <- read_xml(file.path(
   "https://www.wikipathways.org",
   "wikipathways-assets/pathways",
   pid,
   file_name
))
```

## Extract href

In the original SVG, elements come with hyperlinks that impair bscui
interactions. That's why they were removed from the SVG itself and put in
a table for further use with bscui functions.

```{r}
namespace_element_nodes <- xml_find_all(
   svg, "//namespace::*[name()='']/parent::*"
)
xml_attr(namespace_element_nodes, "xmlns") <- NULL
a <- xml_find_all(svg, "//a")
info <- tibble(
   id = xml_attr(a, "id"),
   name = xml_attr(a, "name"),
   href = xml_attr(a, "href"),
   svg_class = xml_attr(a, "class")
) |> 
   mutate(
      category = case_when(
         str_detect(svg_class, "GeneProduct") ~ "GeneProduct",
         str_detect(svg_class, "Metabolite") ~ "Metabolite",
         TRUE ~ as.character(NA)
      ),
      ensembl = strsplit(svg_class, split = " ") |> 
         lapply(function(x){
            grep("^Ensembl_", x, value=TRUE) |> 
               str_remove("Ensembl_") |> 
               paste(collapse=", ")
         }) |> 
         unlist(),
      ChEBI = strsplit(svg_class, split = " ") |> 
         lapply(function(x){
            grep("^ChEBI_", x, value=TRUE) |> 
               str_remove("ChEBI_") |> 
               paste(collapse=", ")
         }) |> 
         unlist()
   )
xml_attr(a, "target") <- NULL
xml_attr(namespace_element_nodes[1], "xmlns") <- "http://www.w3.org/2000/svg"
svg <- as.character(svg) |>
   stringr::str_remove_all('xlink:href="https?://[^"]*"') |>
   read_xml()
```

Here is the pre-processed SVG in a bscui figure:

```{r}
bscui(svg)
```

```{r, echo=FALSE}
write_xml(svg, file.path(data_dir, str_replace(file_name, ".svg$", ".svg.gz")))
```

```{r, echo=FALSE}
readr::write_tsv(
   info,
   file=file.path(data_dir, str_replace(file_name, ".svg$", ".txt.gz"))
)
```

