---
title: "Preparing SVG: examples, tips and tricks"
package: "bscui (version `r packageVersion('bscui')`)"
output:
   rmarkdown::html_document:
      code_folding: show
      number_sections: yes
      self_contained: true
      theme: cerulean
      toc: yes
      toc_float: yes
      fig_width: 7
      fig_height: 5
editor_options:
  chunk_output_type: console
---


```{r setup, message=FALSE, echo=FALSE, include=FALSE, cache=FALSE}
library(knitr)
opts_chunk$set(
   include=TRUE,
   echo=TRUE,
   message=TRUE,
   warning=TRUE,
   cache=FALSE,
   cache.lazy=FALSE
)
library(here)
library(bscui)
library(xml2)
library(dplyr)
library(readr)
##
data_dir <- here("inst/svg-examples")
```

```{js, echo=FALSE}
function toggleCollapse(id) {
   var content = document.querySelector('#'+id);
   if(content.style.display === 'none' || content.style.display === ''){
      content.style.display = 'block';
   }else{
      content.style.display = 'none';
   }
}
```

```{css, echo=FALSE}
/* Hide the content of the collapsible div by default */
.collapsible {
   display: none;
   border: solid 1px black;
   background-color: #FAFA000F;
   margin: 5px 0px 5px 0px;
   padding: 5px;
}

.collapse-button {
   border: solid 1px black;
   border-radius: 3px;
   background-color: #CFCFCF;
}
```


# Installation and requirements

The following packages will be used.

```{r}
library(bscui)
library(xml2)
library(dplyr)
library(readr)
library(stringr)
```

<!-- Collapsible sessionInfo -->
<button class="collapse-button" onclick="toggleCollapse('sessionInfo')">
Display session info
</button>

::: {#sessionInfo class="collapsible" style="margin-top:10px; margin-bottom:10px;"}

Built on `r Sys.Date()` 

```{r, class.source='fold-hide'}
sessionInfo()
```
:::

# Reading anatomograms

## Original SVG

From: https://github.com/ebi-gene-expression-group/anatomogram

```{r}
file_name <- "homo_sapiens.female.svg"
svg <- read_xml(file.path(
   "https://raw.githubusercontent.com",
   "ebi-gene-expression-group/anatomogram",
   "master/src/svg",
   file_name
))
```

## Extract element titles from the SVG

```{r}
get_element_titles <- function(x){
   id <- xml_attr(x, "id")
   children <- xml_children(x)
   children_titles <- tibble(id=character(), label=character())
   label <- character()
   if(length(children) > 0) for(i in 1:length(children)){
      child <- children[[i]]
      if(xml_name(child) == "title"){
         label <- xml_attr(child, "id")
      }else{
         children_titles <- children_titles |>
            bind_rows(get_element_titles(child))
      }
   }
   toRet <- tibble(id = id, label = label)  |>
      bind_rows(children_titles) |>
      filter(!is.na(id))
   return(toRet)
}
element_titles <- get_element_titles(svg)
```

```{r}
head(element_titles)
```

## Remove title elements from SVG

```{r}
namespace_element_nodes <- xml_find_all(
   svg, "//namespace::*[name()='']/parent::*"
)
xml_attr(namespace_element_nodes, "xmlns") <- NULL
titles <- xml_find_all(svg, "//title")
for(to_remove in titles){
   xml_remove(to_remove)
}
xml_attr(namespace_element_nodes[1], "xmlns") <- "http://www.w3.org/2000/svg"
```

```{r, echo=FALSE}
write_xml(svg, file.path(data_dir, str_replace(file_name, ".svg$", ".svg.gz")))
write_tsv(
   element_titles,
   file.path(data_dir, str_replace(file_name, ".svg$", ".txt.gz"))
)
```

```{r}
bscui(svg)
```

# Reading swissbiopics

## Read original SVG

From: https://www.swissbiopics.org/

```{r}
file_name <- "Animal_cells.svg"
svg <- read_xml(file.path(
   "https://www.swissbiopics.org/",
   "api/image",
   file_name
))
```

## Remove text elements from SVG

```{r}
namespace_element_nodes <- xml_find_all(
   svg, "//namespace::*[name()='']/parent::*"
)
xml_attr(namespace_element_nodes, "xmlns") <- NULL
titles <- xml_find_all(svg, "//text")
for(to_remove in titles){
   xml_remove(to_remove)
}
xml_attr(namespace_element_nodes[1], "xmlns") <- "http://www.w3.org/2000/svg"
```


```{r, echo=FALSE}
write_xml(svg, file.path(data_dir, str_replace(file_name, ".svg$", ".svg.gz")))
```

```{r}
bscui(svg)
```

## Get cellular location information

```{r}
tmp_file <- tempfile(fileext=".txt.gz")
info <- download.file(
   url = "https://rest.uniprot.org/locations/stream?compressed=true&fields=id%2Cname%2Ccategory%2Cdefinition%2Ccontent%2Cgene_ontologies&format=tsv&query=%28*%29",
   destfile = tmp_file
)
info <- readr::read_tsv(
   tmp_file,
   col_types = strrep("c", 6)
)
```

```{r, echo=FALSE}
file_name <- "uniprot_cellular_locations.txt.gz"
readr::write_tsv(info, file=file.path(data_dir, file_name))
```

# Read wikipathways

## Read original SVG

From: https://www.wikipathways.org/

```{r}
pid <- "WP4657"
# pid <- "WP521"
file_name <- paste0(pid, ".svg")
svg <- read_xml(file.path(
   "https://www.wikipathways.org",
   "wikipathways-assets/pathways",
   pid,
   file_name
))
```

## Extract href

```{r}
namespace_element_nodes <- xml_find_all(
   svg, "//namespace::*[name()='']/parent::*"
)
xml_attr(namespace_element_nodes, "xmlns") <- NULL
a <- xml_find_all(svg, "//a")
href <- tibble(
   id = xml_attr(a, "id"),
   href = xml_attr(a, "href")
)
xml_attr(a, "target") <- NULL
xml_attr(namespace_element_nodes[1], "xmlns") <- "http://www.w3.org/2000/svg"
svg <- as.character(svg) |>
   stringr::str_remove_all('xlink:href="https?://[^"]*"') |>
   read_xml()
```


```{r, echo=FALSE}
write_xml(svg, file.path(data_dir, str_replace(file_name, ".svg$", ".svg.gz")))
```



```{r}
bscui(svg)
```
