---
title: "Introduction to bscui"
package: "bscui (version `r packageVersion('bscui')`)"
output:
   rmarkdown::html_document:
      code_folding: show
      number_sections: yes
      self_contained: true
      theme: cerulean
      toc: yes
      toc_float: yes
      fig_width: 7
      fig_height: 5
vignette: >
   %\VignetteIndexEntry{bscui}
   %\VignetteEncoding{UTF-8}
   %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, message=FALSE, echo=FALSE, include=FALSE, cache=FALSE}
library(knitr)
opts_chunk$set(
   include=TRUE,
   echo=TRUE,
   message=TRUE,
   warning=TRUE,
   cache=FALSE,
   cache.lazy=FALSE
)
library(bscui)
library(dplyr)
library(xml2)
library(dplyr)
library(readr)
library(stringr)
library(glue)
library(DT)
```

```{js, echo=FALSE}
function toggleCollapse(id) {
   var content = document.querySelector('#'+id);
   if(content.style.display === 'none' || content.style.display === ''){
      content.style.display = 'block';
   }else{
      content.style.display = 'none';
   }
}
```

```{css, echo=FALSE}
/* Hide the content of the collapsible div by default */
.collapsible {
   display: none;
   border: solid 1px black;
   background-color: #FAFA000F;
   margin: 5px 0px 5px 0px;
   padding: 5px;
}

.collapse-button {
   border: solid 1px black;
   border-radius: 3px;
   background-color: #CFCFCF;
   margin-top: 10px;
   margin-bottom: 20px;
}
```

The aim of the 'bscui' R package is to use any SVG image as an interactive figure
on which contextual information can be displayed.
These figures can be integrated
in [R Markdown](https://rmarkdown.rstudio.com/)/[Quarto](https://quarto.org/)
documents or in [Shiny](https://shiny.posit.co/) applications.

Here are the main features of 'bscui' figures:

   - Interactive view: pan, zoom in/out
   - Update the style (e.g., color, stroke, opacity...) of SVG elements
   - Update the attribute (e.g., path) of SVG elements
   - Add or remove SVG elements
   - Export the figure in SVG or PNG format
   - Display contextual information when hovering over SVG elements
   - Select one or multiple SVG elements (mainly used with Shiny)
   - Click on elements (mainly used with Shiny)

The main motivation behind the development of this package was to be able
to leverage SVG images kindly shared in public knowledge resources
(e.g.,
[EBI anatomograms](https://ebi-gene-expression-group.github.io/anatomogram/),
[SwissBioPics](https://www.swissbiopics.org/),
or [wikipathways](https://www.wikipathways.org/)
)
to display and browse information in the relevant contexts.

The package comes with several pre-processed
examples used to demonstrate the available features, as shown in this document.

# Installation and requirements

The 'bscui' R package is currently available
in [github](https://github.com/patzaw/bscui/).

```{r, eval = FALSE}
## Dependencies
install.packages("htmlwidgets")
## Install from github
devtools::install_github("patzaw/bscui")
```

Hopefully 'bscui' will be available soon on [CRAN](https://cran.r-project.org/).

## Vignette requirements

The following packages are not dependencies of 'bscui' strictly speaking but 
are required to run code examples discussed in this document.

```{r, eval = FALSE}
library(xml2)
library(dplyr)
library(readr)
library(stringr)
library(glue)
```


<!-- Collapsible sessionInfo -->
<button class="collapse-button" onclick="toggleCollapse('sessionInfo')">
Display session info
</button>

::: {#sessionInfo class="collapsible" style="margin-top:10px; margin-bottom:10px;"}

Built on `r Sys.Date()` 

```{r, class.source='fold-hide'}
sessionInfo()
```
:::

- [xml2](https://cran.r-project.org/package=xml2) is used to read SVG files
- [readr](https://cran.r-project.org/package=readr) and
[dplyr](https://cran.r-project.org/package=readr)
are used to read and manipulate tables with figure contextual information
- [stringr](https://cran.r-project.org/package=stringr) and
[glue](https://cran.r-project.org/package=glue) are used to facilitate
the creation of html tags

# Simple use in R session or static documents

The main function of the package is `bscui()`. It takes as main argument
a character string with SVG code.
The use of [xml2](https://cran.r-project.org/package=xml2) to read SVG files is
not mandatory but it can be useful to validate the SVG and manipulate it before
displaying it.

## Animal cells example

This example relies on a figure of animal cells taken
from [SwissBioPics](https://www.swissbiopics.org/name/Animal_cell).

```{r}
svg <- xml2::read_xml(system.file(
   "svg-examples", "Animal_cells.svg.gz",
   package="bscui"
))
```

The simplest way to display the figure is shown below.

```{r}
figure <- bscui(svg)
figure
```

The figure can be grabbed with mouse and enlarged or shrunk using
the mouse wheel. Clicking on the button at the top-left corner of
the figure displays a menu with various functions, including
resetting the view and exporting the figure in SVG or PNG format.
Several configuration choices are made by default but can be changed,
such as the zoom range or the width of the menu.

## Defining UI elements

User Interface (UI) elements are provided as a data frame with the following
columns:

   - **id**: the element identifier
   - **ui_type**: either "selectable" (several elements can be selected),
   "button" (action will be triggered by clicking), "none" (no action on click)
   - **title**: a description of the element to display the mouse hovers
   over the element
   
Information about the different part of a cell were taken
from [UniProt](https://www.uniprot.org/locations).

```{r}
info <- readr::read_tsv(system.file(
   "svg-examples", "uniprot_cellular_locations.txt.gz",
   package="bscui"
), col_types=strrep("c", 6)) |> 
   mutate(id = str_remove(`Subcellular location ID`, "-"))
```

The title feature can take simple text or a valid html element as shown below. 

```{r}
ui_elements <- info |> 
   mutate(
      ui_type = "selectable",
      title = glue(
         '<div style="width:300px; height:100px; overflow:auto; padding:5px;',
         'font-size:75%;',
         'border:black 1px solid; background:#FFFFF0AA;">',
         "<strong>{Name}</strong>: {Description}",
         "</div>",
         .sep=" "
      )
   ) |>
   select(id, ui_type, title)
figure <- figure |> 
   set_bscui_ui_elements(ui_elements)
figure
```

## Setting element styles

The style of the elements can be changed with the `set_bscui_styles()` function.
This function can be called several times to change different elements or
different options.

```{r}
figure <- figure |> 
   set_bscui_styles(
      bind_rows(
         info |>
            filter(Name == "Cytosol") |>
            mutate(fill = "#FF7F7F"),
         info |>
            filter(Name == "Nucleoplasm") |>
            mutate(fill = "#7F7FFF")
      ) |> 
         select(
            id, fill
         )
   ) |> 
   set_bscui_styles(
      info |>
         filter(Name == "Endosome") |>
         mutate(stroke = "yellow", strokeWidth = "2px") |> 
         select(id, stroke, strokeWidth)
   )
figure
```

## Setting element attributes

Element attributes can also be changed with
the `set_bscui_attributes()` function. The example below shows how to focus
on the nucleus and its content starting from the original figure, by setting
the "display" attribute of all other elements to "none". Also the size of the
nucleolus is slightly increased.

```{r}
nucleus_part <- c(
   "SL0191", "SL0190", "SL0182", "SL0188", "SL0494", "SL0180",
   "SL0031", "SL0465", "SL0127", "SL0186"
)
figure <- figure |>
   set_bscui_attributes(
      info |>
         filter(
            !id %in% nucleus_part
         ) |>
         mutate(display="none") |>
         select(id, display)
   ) |> 
   set_bscui_attributes(tibble(id="sib_copyright", display="none")) |>
   set_bscui_attributes(tibble(id="SL0188", transform="scale(1.8 1.8)")) |> 
   set_bscui_attributes(
      tibble(id="SL0188", transform="translate(-237 -202)"),
      append=TRUE
   )
figure |> 
   set_bscui_options(show_menu=FALSE, zoom_min=1, zoom_max=1, clip=TRUE)
```

In the example above the `set_bscui_options()` function is used to hide the
menu and to disable view changes (zoom and pan). Other options can be changed
with this function.

# Shiny applications

## Reacting

## Updating

